ROBUST BAYESIAN RELIABILITY FOR COMPLEX SYSTEMS UNDER PRIOR-DATA CONFLICT

### 2016-07-29 asce-paper.tex
finishing work on examples and examples section

\title{Notes for ISIPTA poster}
\author{Gero Walter, Frank P.A. Coolen, Simme Douwe Flapper}

\begin{document}
\title{ROBUST BAYESIAN RELIABILITY FOR COMPLEX SYSTEMS\\ or \ldots}

\author{
Gero Walter%
\thanks{
School of Industrial Engineering,
Eindhoven University of Technology, Eindhoven, The Netherlands.
E-mail: g.walter@tue.nl.},
\ Ph.D.
\\
and
Frank P.A. Coolen%
\thanks{
Department of Mathematical Sciences,
Durham University, Durham, United Kingdom.
E-mail: frank.coolen@durham.ac.uk.},
\ Ph.D.
\\
and ???
}

@@

Invited for special issue ``The treatment of uncertainty in risk and reliability modelling and decision making''
(special issue code SI022A), guest editors Luca Podolfillini, Bruno Sudret, Enrico Zio,
in ASCE-ASME Journal of Risk and Uncertainty in Engineering Systems: Part A (Civil Engineering)

***``For most ASCE journals, the maximum number of words and word-equivalents is 10,000 for technical papers''

***Abstract must be 150--175 words long (this has 175), may not contain references or mathematics.

\begin{abstract}
In reliability engineering, data about failure events is often scarce.
To arrive at meaningful estimates for the reliability of a system,
it is therefore often necessary to also include expert information in the analysis,
which is straightforward in the Bayesian approach by using an informative prior distribution.
%
A problem called prior-data conflict then can arise:
observed data seem very surprising from the viewpoint of the prior,
i.e., information from data is in conflict with prior assumptions.
Models based on conjugate priors can be insensitive to prior-data conflict,
in the sense that the spread of the posterior distribution does not increase in case of such a conflict,
thus conveying a false sense of certainty.
%
We present an approach to mitigate this issue, by considering sets of prior distributions
to model limited knowledge on Weibull distributed component lifetimes,
treating systems with arbitary layout using the survival signature.
Our approach can be seen as a robust Bayesian procedure or imprecise probability method
that reflects surprisingly early or late component failures
by wider system reliability bounds.
%We show how surprisingly early or late component failures
%result in wider system reliability bounds.
%
%and study how surprisingly early or late component failures
%affect the prediction of the reliability of a system with arbitrary layout,
%making use of survival signature to characterise the system under study.
%Our approach can be seen as a robust Bayesian procedure or imprecise probability method
%that appropriately reflects surprising data in the posterior system reliability function.
\end{abstract}
% by communicating that we can quantify the reliability of a system quite precisely when in fact we cannot.

@@

These prior assumptions for the four component types are visualized in Figure~\ref{fig:brake-comppriors},
showing the sets of reliability functions corresponding to the prior predictive density \eqref{eq:tpriopred}.
The figure thus displays the bounds for the probability that a single component,
having been put under risk at time $0$, will have failed by time $t$.
The top left graph in Figure~\ref{fig:brake-sysrels} shows what the prior assumptions on components
signify for the system, depicting the prior bounds for the system reliability %before the system has started
on a scale of time elapsed since system startup.
For example, the prior probability of the system to survive until time $10$ is between $***$ and $***$.
%calender time scale, where $t = 0$ is the time when the system starts to run.

@@

\begin{figure}
\includegraphics[width=\textwidth]{fig5a}
\caption{Sets of posterior predictive reliability functions for the four component types
for observations in line with prior expectations.}
\label{fig:comppost-1}
\end{figure}

We see that the graphs for M and H do not change dramatically,
as there is only one component of each in the system to learn from.
For C, the bounds have considerably narrowed,
showing the effect of having seen the four observations
$t_1^\text{C} = 6$, $t_2^\text{C} = 7$, $t_3^\text{C} = t_4^\text{C} = 8^+$.
The bounds for P have not narrowed as much;
this is due to the two right-censored observations $t_3^\text{P} = t_4^\text{P} = 8^+$;
from the prior assumptions, surviving past time 8 is already quite unusual.

@@

For the second case, with $t_1^\text{C} = 1$, $t_2^\text{C} = 2$, $t_1^\text{P} = 0.5$, $t_2^\text{P} = 1$
and $\tnow = 2$, we assume to observe surprisingly early failures;
the corresponding posterior predictive component distributions are given in Figure~\ref{fig:comppost-2}.
***some discussion***

\begin{figure}
\includegraphics[width=\textwidth]{fig6a}
\caption{Sets of posterior predictive reliability functions for the four component types
for surprisingly early failures.}
\label{fig:comppost-2}
\end{figure}

@@

For the third case, we assume to observe surprisingly late failures,
namely $t_1^\text{C} = 11$, $t_2^\text{C} = 12$, $t_1^\text{P} = 8$, $t_2^\text{P} = 9$, and $\tnow = 12$;
the corresponding posterior predictive component distributions are given in Figure~\ref{fig:comppost-3}.
***some discussion***

\begin{figure}
\includegraphics[width=\textwidth]{fig7a}
\caption{Sets of posterior predictive reliability functions for the four component types
for surprisingly late failures.}
\label{fig:comppost-3}
\end{figure}

@@

Figure~\ref{fig:brake-sysrels} depicts the set of prior system reliability functions,
together with the sets of posterior system reliability function for the three cases.
***some discussion***

\begin{figure}
\includegraphics[width=\textwidth]{fig8}
\caption{Sets of prior and posterior system reliability functions for the three cases.*** }
\label{fig:brake-sysrels}
\end{figure}

Figure~\ref{fig:brake-sysrels-imprecision} shows $\uRsys(t) - \lRsys(t)$ ***

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fig9}
\caption{Imprecision of prior and posterior system reliability sets.*** }
\label{fig:brake-sysrels-imprecision}
\end{figure}

>>>>

\begin{document}
\title{ROBUST BAYESIAN RELIABILITY FOR COMPLEX SYSTEMS UNDER PRIOR-DATA CONFLICT}

\author{
Gero Walter%
\thanks{
School of Industrial Engineering,
Eindhoven University of Technology, Eindhoven, The Netherlands.
E-mail: g.walter@tue.nl. Corresponding author.},
\ Ph.D.
\\
and
Frank P.A. Coolen%
\thanks{
Department of Mathematical Sciences,
Durham University, Durham, United Kingdom.
E-mail: frank.coolen@durham.ac.uk.},
\ Ph.D.
}

@@

%Invited for special issue ``The treatment of uncertainty in risk and reliability modelling and decision making''
%(special issue code SI022A), guest editors Luca Podolfillini, Bruno Sudret, Enrico Zio,
%in ASCE-ASME Journal of Risk and Uncertainty in Engineering Systems: Part A (Civil Engineering)
%
%***``For most ASCE journals, the maximum number of words and word-equivalents is 10,000 for technical papers''
%***Abstract must be 150--175 words long (this has 175), may not contain references or mathematics.
%
\begin{abstract}
In reliability engineering, data about failure events is often scarce.
To arrive at meaningful estimates for the reliability of a system,
it is therefore often necessary to also include expert information in the analysis,
which is straightforward in the Bayesian approach by using an informative prior distribution.
%
A problem called prior-data conflict then can arise:
observed data seem very surprising from the viewpoint of the prior,
i.e., information from data is in conflict with prior assumptions.
Models based on conjugate priors can be insensitive to prior-data conflict,
in the sense that the spread of the posterior distribution does not increase in case of such a conflict,
thus conveying a false sense of certainty.
%
We present an approach to mitigate this issue, by considering sets of prior distributions
to model limited knowledge on Weibull distributed component lifetimes,
treating systems with arbitary layout using the survival signature.
Our approach can be seen as a robust Bayesian procedure or imprecise probability method
that reflects surprisingly early or late component failures
by wider system reliability bounds.
%We show how surprisingly early or late component failures
%result in wider system reliability bounds.
%
%and study how surprisingly early or late component failures
%affect the prediction of the reliability of a system with arbitrary layout,
%making use of survival signature to characterise the system under study.
%Our approach can be seen as a robust Bayesian procedure or imprecise probability method
%that appropriately reflects surprising data in the posterior system reliability function.
\end{abstract}
% by communicating that we can quantify the reliability of a system quite precisely when in fact we cannot.

@@

These prior assumptions for the four component types are visualized in Figure~\ref{fig:brake-comppriors},
showing the sets of reliability functions corresponding to the prior predictive density \eqref{eq:tpriopred}.
The figure thus displays the bounds for the probability that a single component,
having been put under risk at time $0$, will have failed by time $t$.
The top left graph in Figure~\ref{fig:brake-sysrels} shows what the prior assumptions on components
signify for the system, depicting the prior bounds for the system reliability %before the system has started
on a scale of time elapsed since system startup.
For example, the prior probability of the system to survive until time $10$ is between $0.03$\% and $6.91$\%.
%calender time scale, where $t = 0$ is the time when the system starts to run.

@@

\begin{figure}
\includegraphics[width=\textwidth]{fig5a}
\caption{Sets of posterior predictive reliability functions for the four component types
for observations in line with prior expectations (case~1).}
\label{fig:comppost-1}
\end{figure}

We see that the graphs for M and H do not change dramatically,
as there is only one component of each in the system to learn from.
For C, the bounds have considerably narrowed,
showing the effect of having seen the four observations
$t_1^\text{C} = 6$, $t_2^\text{C} = 7$, $t_3^\text{C} = t_4^\text{C} = 8^+$.
The bounds for P have not narrowed as much;
this is due to the two right-censored observations $t_3^\text{P} = t_4^\text{P} = 8^+$;
from the viewpoint of the prior, surviving past time 8 is already quite unusual.

@@

For the second case, with $t_1^\text{C} = 1$, $t_2^\text{C} = 2$, $t_1^\text{P} = 0.25$, $t_2^\text{P} = 0.5$
and $\tnow = 2$ (so $t_1^\text{M} = t_1^\text{H} = t_3^\text{C} = t_4^\text{C} = t_3^\text{P} = t_4^\text{P} = 2^+$),
we assume to observe surprisingly early failures;
the corresponding posterior predictive component distributions are given in Figure~\ref{fig:comppost-2}.
%
\begin{figure}
\includegraphics[width=\textwidth]{fig6a}
\caption{Sets of posterior predictive reliability functions for the four component types
for surprisingly early failures (case~2).}
\label{fig:comppost-2}
\end{figure}
%
Having observed the system only until $t = 2$, our data is not very informative,
such that prior and posterior predictive reliability bounds are very similar.
For C, the effect of the early failures is however still visible,
and posterior imprecision, i.e., the range between lower and upper posterior bound, is notably larger
as compared to prior imprecision, and substantially larger that posterior imprecision in case~1.
The effect for P is less pronounced,
mainly because observing $\vec{t}^\text{P} = (0.25, 0.5, 2^+, 2^+)$ for P is not as extreme
as observing $\vec{t}^\text{C} = (1, 2, 2^+, 2^+)$ for C.

@@

For the third case, we assume to observe surprisingly late failures,
namely $t_1^\text{C} = 11$, $t_2^\text{C} = 12$, $t_1^\text{P} = 8$, $t_2^\text{P} = 9$,
and $\tnow = 12$ (so $t_1^\text{M} = t_1^\text{H} = t_3^\text{C} = t_4^\text{C} = t_3^\text{P} = t_4^\text{P} = 12^+$);
the corresponding posterior predictive component distributions are given in Figure~\ref{fig:comppost-3}.
%
\begin{figure}
\includegraphics[width=\textwidth]{fig7a}
\caption{Sets of posterior predictive reliability functions for the four component types
for surprisingly late failures (case~3).}
\label{fig:comppost-3}
\end{figure}
%
Having observed the system for a much longer time than in case 2,
our data is much more informative,
resulting in considerable differences between prior and posterior bounds. %predictive reliability
The effect of these surprisingly late failures is most prominent for P,
with a set considerably shifted to the right and having very wide posterior bounds.
The posterior set for C also indicates that,
after having seen these late failures, we expect type C components to fail much later.
This effect is also visible for M and H,
but is weaker for them as there is only one component of each in the system.

@@

\begin{figure}
\includegraphics[width=\textwidth]{fig8}
\caption{Sets of prior and posterior system reliability functions for the three cases
on a time showing time elapsed since system startup.}
\label{fig:brake-sysrels}
\end{figure}
%
Figure~\ref{fig:brake-sysrels} depicts the set of prior system reliability functions,
together with the sets of posterior system reliability function for the three cases,
on a scale of elapsed time since system startup.
Due to this time scale, posterior system reliability is 1 at $\tnow$,
as we know that the system has survived until $\tnow = 8, 2, 12$ in each of the three cases.
For all three cases, the posterior bounds drop faster after $\tnow$ than the prior bounds drop after $t = 0$
since the components in the system have aged until $\tnow$ and so are expected to fail sooner.
In the case of surprisingly early failures, posterior bounds are mostly within prior bounds,
this is due to $\tnow$ being close to $0$ and weakly informative data in this scenario;
the posterior bounds are nevertheless wider than those for case~1;
posterior bounds are widest for case~3.
%
\begin{figure}
%\centering
\includegraphics[width=\textwidth]{fig9}
\caption{Imprecision of prior and posterior system reliability sets.}
\label{fig:brake-sysrels-imprecision}
\end{figure}
%
This most visible in Figure~\ref{fig:brake-sysrels-imprecision},
which shows $\uRsys(t) - \lRsys(t)$, the difference between upper and lower bound
of prior and posterior system reliability.
The left panel shows imprecision on the scale of elapsed time since system startup like in Figure~\ref{fig:brake-sysrels};
the right panel instead on the scale of prospective time,
indicating how far in the future periods are for which estimation of system reliability is most uncertain.
We see that posterior imprecision is indeed considerably lower in case~1,
where we had observed failures more or less like expected.
On a prospective timescale, we see that periods of heightened uncertainty are closer to the present for the posteriors,
while uncertainty is considerably reduced for periods further in the future.
%This has interesting indications for maintenance planning,
%where uncertainty about system failure in the near future
